### 2024-08-07 paper/paper.qmd
Data commit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Rahma Binth Mohammad
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)

total_ex_cpi <- read_csv(here::here("data/analysis_data/total_ex_cpi.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
```{r}
#| fig-cap: Linear model
#| echo: false
#| warning: false
#| message: false

total_ex_cpi |>
  ggplot(aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.5, aes(group = 1)) + 
  labs(
    title = "Total Expenditure vs All-Items CPI",
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  theme_minimal() +
  scale_color_viridis_d()
```

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

### Simple linear regression

Define $y_i$ as the total expenditure in Canadian dollars in year i. Then $cpi_i$ is the annual Consumer Price Index (CPI) for year i

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{cpi}_i + \epsilon\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Where

### Multiple linear regression

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{cpi}_i + \beta_2 \mbox{income}_i+ \epsilon\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

```{r}
#| label: fig-2
#| fig-cap: Linear model
#| echo: false
#| warning: false
#| message: false
library(scales)

total_ex_cpi |>
  ggplot(aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = y ~ x
  ) +
  labs(
    title = "Total Expenditure vs All-Items CPI",
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_viridis_d() +
  theme_minimal()
```

```{r}
#| label: fig-3
#| fig-cap: Linear model
#| echo: false
#| warning: false
#| message: false
library(scales)

total_ex_cpi |>
  ggplot(aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = y ~ x
  ) +
  labs(
    title = "Total Expenditure vs All-Items CPI",
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_viridis_d() +
  theme_minimal()
```
# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of "
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model,
    "Second model" = second_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check