### 2024-08-08 paper/paper.qmd
Update graphs and add to paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Rahma Binth Mohammad
thanks: "Code and data are available at: https://github.com/rahmabinth/consumer-behaviour"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(patchwork)
spending <-  read_csv(here::here("data/analysis_data/spending.csv"))
cpi <-  read_csv(here::here("data/analysis_data/cpi.csv"))
income <- read_csv(here::here("data/analysis_data/income.csv"))

total_ex_cpi <- read_csv(here::here("data/analysis_data/total_ex_cpi.csv"))

```

# Introduction

# Data {#sec-data}

```{r}
#| label: fig-spending
#| fig-cap: Household spending per year
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false
#Plot spending categories for the years
spending_colours <- c(
  "Total expenditure" = "goldenrod3", 
  "Food" = "purple",
  "Shelter" = "green",
  "Recreation" = "blue")
ggplot(spending, aes(x = Year, y = VALUE, color = expenditure_categories, group = expenditure_categories)) +
  geom_line() +
  geom_point() +
  labs(

    x = "Year",
    y = "Expenditure Value ($)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = spending_colours, name = NULL) +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2)) +
  scale_y_continuous(
      labels = scales::comma
    )
```
From the spending dataset, we chose to select the values for all food expenditures, all shelter expenditures, recreation expenditures, and total expenditures. This choice was made because these are principal expenditures for everyone. Recreation was added to this mix, to identify how consumers behave when inflation is high.

```{r}
#| label: fig-cpi
#| fig-cap: Consumer Price Index (CPI) per Year
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false

cpi_colours <- c(
  "All-items" = "goldenrod3", 
  "Food" = "purple",
  "Shelter" = "green",
  "Recreation" = "blue")

ggplot(cpi, aes(x = Year, y = VALUE, color = product_groups, group = product_groups)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Year",
    y = "CPI",
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = cpi_colours, name = NULL) +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2)) +
  scale_y_continuous(
    breaks = seq(100, 160, by = 10),
    limits = c(100, 160)
  )

```

From the CPI dataset, we chose to look at the annual average to keep it consistent with the annual average of expenditures. The variables

```{r}
#| label: fig-income
#| fig-cap: Number of Income Earners
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false
#Plot the number of income earners against the year
ggplot(income, aes(x = Year, y = VALUE)) +
  geom_line(colour = "darkred") +
  geom_point(colour = "darkred") +
  labs(
    x = "Year",
    y = "Income Earners"
  ) +
  theme_minimal() +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2)) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(9600000, 11250000)
  )

```

From the income dataset, we chose to isolate the values for the category “all persons with income”. This indicate all tax fillers and dependants that earn income from all sources. This was done to keep our analysis consistent.

```{r}
#| label: fig-total_ex_cpi
#| fig-cap: CPI and total expenditure over the year
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false

ggplot(total_ex_cpi, aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.5, aes(group = 1)) + # Connect the points with a line
  labs(
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_x_continuous(
    limits = c(115, 145)) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(75000, 100000)
  )
```

# Model

The goal of our model is to forecasting the total expenditure in a household in a year, based solely on a the Consumer Price Index, and the number of people in Ontario earning an income. We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

### Simple linear regression
The model we employ is:

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{cpi}_i + \epsilon\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Where:

* $y_i$ as the total expenditure in Canadian dollars in year i. 
* $cpi_i$ is the annual Consumer Price Index (CPI) for year i

### Multiple linear regression
The model we employ is:

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{cpi}_i + \beta_2 \mbox{income}_i+ \epsilon\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Where:
* $y_i$ as the total expenditure in Canadian dollars in year i. 
* $cpi_i$ is the annual Consumer Price Index (CPI) for year i
* $income_i$ is the number of people in Ontario earning an income

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

```{r}
#| label: fig-2
#| fig-cap: Linear model
#| echo: false
#| warning: false
#| message: false
library(scales)

total_ex_cpi |>
  ggplot(aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = y ~ x
  ) +
  labs(
    title = "Total Expenditure vs All-Items CPI",
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_viridis_d() +
  theme_minimal()
```

```{r}
#| label: fig-3
#| fig-cap: Linear model
#| echo: false
#| warning: false
#| message: false
library(scales)

total_ex_cpi |>
  ggplot(aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = y ~ x
  ) +
  labs(
    title = "Total Expenditure vs All-Items CPI",
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_viridis_d() +
  theme_minimal()
```
# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of "
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model,
    "Second model" = second_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check