### 2024-08-11 paper/paper.qmd
Update model to complete prior and posterior checks

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Rahma Binth Mohammad
thanks: "Code and data are available at: https://github.com/rahmabinth/consumer-behaviour"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(patchwork)
library(rstanarm)

spending <-  read_csv(here::here("data/analysis_data/spending.csv"))

cpi <-  read_csv(here::here("data/analysis_data/cpi.csv"))

income <- read_csv(here::here("data/analysis_data/income.csv"))

total_ex_cpi <- read_csv(here::here("data/analysis_data/total_ex_cpi.csv"))

shelter_ex_cpi <- read_csv(here::here("data/analysis_data/shelter_ex_cpi.csv"))

food_ex_cpi <- read_csv(here::here("data/analysis_data/food_ex_cpi.csv"))

recreation_ex_cpi <- read_csv(here::here("data/analysis_data/recreation_ex_cpi.csv"))

spending_n_income <- read_csv(here::here("data/analysis_data/spending_n_income.csv"))

```

# Introduction

Consumption is the biggest indicator of a person’s needs and desires. Our interest is to inquire into whether tracking consumption, along with income and prices of items through inflation, will tell us a better story about a household’s financial behaviour and broader economic trends. By assessing these variables, we can uncover how inflation influences spending habits and whether an increase or decrease in income can mitigate or exacerbate the effects of rising or falling prices. The relationship between these variables can provide information in regards improving economic stability and effective social/governmental assistance to support households in managing their spending in the face of inflationary pressures. Policymakers can use this information to employ strategies for the wellbeing of the state.

Recently, inflation has become serious problem. It diminishes purchasing power, which is a problem at the individual level but also state level. On an individual level, it affects costs of living, impeding the ability to live luxuriously. On a state level, it can lead to increased costs for goods and services, reducing the competitiveness of a state’s exports and making it harder for businesses to thrive on a domestic and global market. Thus, it is crucial states manage inflation as it will strength consumer and investor confidence, leading to economic stability and growth.

This study focuses specifically on the total expenditure, household income, and the inflationary changes over a period of several years.

We used a simple linear regression model and a multiple linear regression for our analysis. For the simple linear model, we used the the change in prices over several years against the average household spending during these years. For the multiple regression model, we used the average household income, along with the change in the prices to predict household spending. 
Our estimand is the

Based on our model and analysis, we found that

This paper was designed to communicate our model and analysis in a structured manner.
@sec-data will provide the datasets used for this analysis. @sec-model will explore the simple and multiple linear regression models used for this analysis, along with the justification for the use of such models. @sec-results will provide the results from that analysis. A discussion of the results and the weaknesses and biases of this type of analysis will be provided in @sec-discussion.

# Data {#sec-data}

## Methodology
The language and environment used for this analysis is R [@citeR], alongside the tidyverse (@tidyverse), rstanarm (@rstanarm), cansim [@cansim], ggplot2 [@ggplot2], scales [@scales], modelsummary [@modelsummary], and marginaleffects [@marginaleffects] packages.

We downloaded the data directly from Statistics Canada using the cansim package. We cleaned it, identified missing values and wrote csv files to make it easily accessible using R studio.

## Sources and Variables {#sec-sources_variables}
A good indicator of inflation is the Consumer Price Index (CPI), as it measures the change in price using a fixed basket of goods and services. Statistics Canada is a government agency that collects information on Canada’s economy (cite). Thus, it has the data that provides the CPI for numerous years. Statistics Canada also provides us with other datasets such as the average household income and the average household expenditures, in a year for all the provinces. We have chosen to look at Ontario solely, as we hold firmly Ontario behaves very differently due to its high population in relation to the other provinces. The data for Canada as a whole is complicated as there are many missing values for some provinces. Our analysis and the data provided by Statistics Canada makes it possible to reproduce our model for different provinces using the github repository.  While there are other datasets available, these three datasets are provided by the same source and has data for provinces across Canada. The expenditure and cpi data also provides the information for three additional specific categories, food, shelter, and recreation, making it easier to isolate and answer our research question. @sec-sources_variables will illustrate the visualised data.

This paper’s objective is to address the following questions using data collected from a government census: (1) Does changes in prices affect expenditure? (2) Does changed in prices affect expenditure as income changes?

### Spending Data

`spending.csv`

This dataset contains the average total expenditures of the Ontario resident for the years 2010 to 2021, excluding years 2018 and 2020. It also contains the breakdown of the expenditures, food purchased (from restaurant and stores), shelter (rented and owned), utilities such as water, fuel, and electricity for the principal accommodation, clothing, transportation (private and public), healthcare, recreational goods and services.

From this dataset, we chose to select the values for all food expenditures, all shelter expenditures, recreation expenditures, and total expenditures. This choice was made because these are principal expenditures for everyone. Recreation was added to this mix, to identify how consumers behave when inflation is high. @fig-food_n_recreation indicates this relationship. In @fig-spending, we observe a steady increase in recreation expenditures, while food, shelter, and overall expenditures show a more rapid growth. Note for simplicity, we did not include the other categories used to calculate total expenditures in the original dataset. This is the reason the total expenditure is significantly higher than the other categories. For all four of the categories, the mean and median values were close to each other, indicating the data is not skewed and did not have many outliers. However, the standard deviation for shelter and total expenditures were higher which indicates more money was spent and thus, more variation over the years. The only variable we use for our models from this dataset is `total expenditure`. However, the model can be used for the other categories. The model is made to be easily reproducible and can be found in the github link provided on page 1.

```{r}
#| label: fig-spending
#| fig-cap: Household spending per year
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false

#Plot spending categories for the years
spending_colours <- c(
  "Total expenditure" = "goldenrod3", 
  "Food" = "purple",
  "Shelter" = "darkgreen",
  "Recreation" = "blue")
ggplot(spending, aes(x = Year, y = VALUE, color = expenditure_categories, group = expenditure_categories)) +
  geom_line() +
  geom_point() +
  labs(

    x = "Year",
    y = "Expenditure Value ($)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = spending_colours, name = NULL) +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2)) +
  scale_y_continuous(
      labels = scales::comma
    )
```

### Inflation Data

`cpi.csv`

The dataset provides CPI for every year. It is not seasonally adjusted for changes as it is an annual average. Thus, it is actual price paid by consumers and not adjusted for seasonal fluctuations. The base year for the CPI calculation is 2002 where CPI is 100. Thus, if our CPI is 120 for a year, it indicates a 20% increase since 2002. This value is useful as an indicator for inflation, as it provides the increase and comparison between years, rather than a absolute price values.

From this dataset, we chose to look at the annual average to keep it consistent with the annual average of expenditures. The data provides cpi for specific categories, as well as all-items. We chose to select cpi for all-items, food, shelter, and recreations, for comparison with the household spending. The prices for food and shelter increased almost 35% in the 10 years illustrated in @fig-cpi. However, the prices for recreation saw a decrease in 2020 and 2021, which indicates a relation to COVID-19. Our models will use the ‘all-items’ CPI to ensure consistency.

```{r}
#| label: fig-cpi
#| fig-cap: Consumer Price Index (CPI) per Year
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false

cpi_colours <- c(
  "All-items" = "goldenrod3", 
  "Food" = "purple",
  "Shelter" = "darkgreen",
  "Recreation" = "blue")

ggplot(cpi, aes(x = Year, y = VALUE, color = product_groups, group = product_groups)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Year",
    y = "CPI",
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = cpi_colours, name = NULL) +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2)) +
  scale_y_continuous(
    breaks = seq(100, 160, by = 10),
    limits = c(100, 160)
  )

```

### Income Data

`income.csv`

Throughout the province of Ontario, there exits people in many different financial situations. Statistics Canada provides us with this dataset that contains the average number of people in an income category. For example, “persons with income under $5000", “persons with income of $5000 and over”, “persons with income of $10 000 and over.” Note for 2020, COVID-19 government income support and benefits are included in income estimates.

From this dataset, we chose to isolate the values for the category “all persons with income”. This indicate all tax fillers and dependants that earn income from all sources. This was done to keep our analysis consistent, as we could not separate who was spending. Our model will use the all persons with income variable as well. @fig-income shows a significant increase in the number of income earners. This could be due to a population increase, specifically an increase in tax fillers and tax dependants.

```{r}
#| label: fig-income
#| fig-cap: Number of Income Earners per Year
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false
#Plot the number of income earners against the year
ggplot(income, aes(x = Year, y = VALUE)) +
  geom_line(colour = "darkred") +
  geom_point(colour = "darkred") +
  labs(
    x = "Year",
    y = "Income Earners"
  ) +
  theme_minimal() +
  scale_x_continuous(
    breaks = seq(2010, 2021, by = 2)) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(9600000, 11250000)
  )

```

For our analysis, we have to use all three dataset values and thus, we chose to use data for the years 2010 to 2021, excluding 2018 and 2020, given the data was only available in all given datasets for these years. However our model does predict the behaviour in 2018 and 2020, using the other years’ data. We also created numerous csv files that isolated the specific data. @sec-appendix thoroughly explains the structure of the csv files.

## Measurement

A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.
A discussion about measurement.

```{r}
#| label: fig-total_n_shelter
#| fig-cap: All-items CPI and total expenditure over the years (on the left), shelter CPI and shelter expenditure over the years (on the right)
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false

fig_total <- ggplot(total_ex_cpi, aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.5, aes(group = 1)) +
  labs(
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_x_continuous(
    limits = c(115, 145)) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(75000, 100000)
  )

fig_shelter <- ggplot(shelter_ex_cpi, aes(x = VALUE.x, y = VALUE.y, color = factor(Year))) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.5,  aes(group = 1)) +
  labs(
    x = "Shelter CPI",
    y = "Shelter Expenditure",
    color = "Year"
  ) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_x_continuous(
    limits = c(120, 160)) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(16000, 25000)
  ) +
  theme(
    legend.position = "none"
  )

total_shelter_cpi <- fig_total + fig_shelter +
  plot_layout(ncol = 2)

print(total_shelter_cpi)

```

```{r}
#| label: fig-food_n_recreation
#| fig-cap: Food CPI and food expenditure over the years (on the left), recreation CPI and recreation expenditure over the years (on the right)
#| fig-align: "center"
#| echo: false
#| warning: false
#| message: false

fig_food <- ggplot(food_ex_cpi, aes(x = VALUE.x, y = VALUE.y, color = factor(Year))) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.5,  aes(group = 1)) +
  labs(
    x = "Food CPI",
    y = "Food Expenditure",
    color = "Year"
  ) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_x_continuous(
    limits = c(120, 160)) +
  scale_y_continuous(
    labels = scales::comma,
    limits = c(7000, 11000)
  )

fig_recreation <- ggplot(recreation_ex_cpi, aes(x = VALUE.x, y = VALUE.y, color = factor(Year))) +
  geom_point(size = 1) +
  geom_line(linewidth = 0.5, aes(group = 1)) +
  labs(
    x = "Recreation CPI",
    y = "Recreation Expenditure",
    color = "Year"
  ) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_y_continuous(
    labels = scales::comma
  ) +
  theme(
    legend.position = "none"
  )

food_recreation_cpi <- fig_food + fig_recreation+
  plot_layout(ncol = 2)

print(food_recreation_cpi)

```

# Model {#sec-model}

In this section we build a model to investigate household consumer behaviour.

## Model set-up

The goal of our model is to forecast the total expenditure in a household in a year, based solely on the Consumer Price Index of all-items, and the number of people in Ontario earning an income. We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`. Background details and diagnostics are included in [Appendix -@sec-model-details].

### Simple linear regression
The model we employ is:

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{cpi}_i + \epsilon\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Where:

* $y_i$ is the continuous outcome variable, representing the total expenditure in Canadian dollars in year i. 
* $cpi_i$ is the predictor variable, representing the annual Consumer Price Index (CPI) for year i

### Multiple linear regression
The model we employed is:

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{cpi}_i + \beta_2 \mbox{income}_i+ \epsilon\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

Where:

* $y_i$ is the continuous outcome variable, representing the total expenditure in Canadian dollars in year $i$. 
* $cpi_i$ is the predictor variable, representing the annual Consumer Price Index (CPI) for year $i$
* $income_i$ is the number of people earning an income in year $i$.

## Model justification

From @fig-total_n_shelter and fig-food_n_recreation, we see a positive relationship in expenditure and CPI as the years pass. Implying that despite rising prices, household spending has continued to increase, specifically in the food and shelter categories. Due to this relationship, we anticipate the relationship is close to linear. We assume this relationship because food and shelter are necessities unlike recreational spending.

```{r}
#| label: fig-2
#| fig-cap: Relationship between CPI for all-items and total household expenditure
#| echo: false
#| warning: false
#| message: false
library(scales)

total_ex_cpi |>
  ggplot(aes(x = cpi, y = expenditure, color = factor(Year))) +
  geom_point(size = 1) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = y ~ x
  ) +
  labs(
    x = "All-Items CPI",
    y = "Total Expenditure",
    color = "Year"
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_viridis_d() +
  theme_minimal()
```

```{r}
#| label: fig-3
#| fig-cap: Relationship between CPI for all-items and total household expenditure
#| echo: false
#| warning: false
#| message: false
library(scales)

```

# Results {#sec-results}

Our model results are summarized in @tbl-modelresults. The table compares our two models, each predicting the outcome variable (total expenditure), based on either CPI only, and CPI and the number of income earners.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of "
#| warning: false

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
second_model <-
  readRDS(file = here::here("models/second_model.rds"))

modelsummary::modelsummary(
  list(
    "CPI" = first_model,
    "Add Income Earners" = second_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}

```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {#sec-appendix}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check